### 2018-11-19 cage_effect.Rmd
T7 no 24/7

---
title: "Cage effect"
author: "Francesco Beghini"
date: "November 8, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(phyloseq)
library(magrittr)
library(tidyverse)
library(reshape2)
library(ggsignif)
library(DESeq2)

```

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/otu_table_mc10_w_tax.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
mappingfile$Description %<>% str_replace_all(.,'_','-')
mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
rownames(mappingfile) <- mappingfile$Description

sample_data(gg_topi) <- mappingfile
sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)

colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")

tax_table_filtered <- tax_table(gg_topi) %>%
  as.data.frame %>% 
  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)
```

# How's the dispostion of cages
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
ordwu <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE)
```

```{r}
plot_ordination(gg_topi, ordwu, color="cage")
```

# B-diversity intra cage
```{r, cache = TRUE}
distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distances %<>%
  inner_join(unique(data.frame(sample_data(gg_topi))[,c('id.topo','cage')]), by = c('topo1' = 'id.topo')) %>%
  rename('cage'='cage1') %>%
  left_join(unique(data.frame(sample_data(gg_topi))[,c('id.topo','cage')]), by = c('topo2' = 'id.topo')) %>%
  rename('cage'='cage2')
```

# Intra cage diversity through time
```{r}
  distances %>%
  filter(cage1==cage2) %>%
  filter(timepoint1==timepoint2) %>%
  ggplot() +
  geom_boxplot(aes(cage1, value)) + 
  facet_wrap(timepoint1 ~ .)
```

# Inter cage diversity through time
```{r}
  distances %>%
  filter(cage1!=cage2) %>%
  filter(timepoint1==timepoint2) %>%
  mutate(cage = paste(.$cage1, .$cage2, sep =' vs ')) %>%
  ggplot() +
  geom_boxplot(aes(cage, value)) + 
  facet_wrap(timepoint1 ~ ., ncol = 1) +
  theme(axis.text.x = element_text(angle = 90))
```